### 2022-12-13 paper.tex
retarded fields

\documentclass{article}
\usepackage{amsmath}
\def\p{\partial}
\begin{document}
If we only take the lowest order effects of the motion of mass sources into account and neglect stresses, then we can define two vector fields, $\vec{E}$ and $\vec{B}$, which are similar to the electric and magnetic field. We can define $A_\mu=-\frac{1}{4}c_0\bar{h}_{0 \mu}=(\varphi/c_0,\vec{A})$, then $A_\mu$, which are similar to the electromagnetic 4-potential, will satisfy
\begin{equation}
    \p^\nu\p_\nu A_\mu=-\frac{4\pi G_0}{c_0^2}J_\mu,\quad\p^\mu A_\mu=0,
\end{equation}
where $J_\mu=-c_0T_{0 \mu}/c_0^2=(c_0\rho,\vec{j})=\rho(c_0,\vec{v})$ is 4-momentum density. If we define $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\varepsilon_{\text{G}0}=\frac{1}{4\pi G_0}$, $\mu_{\text{G}0}=\frac{4\pi G_0}{c_0^2}$, then there will be
\begin{equation}\label{maxwell_0}
    \begin{cases}
        \vec{\nabla}\cdot(\varepsilon_{\text{G}0}\vec{E})=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times(\mu_{\text{G}0}^{-1}\vec{B})=\vec{j}+\frac{\p}{\p t}(\varepsilon_{\text{G}0}\vec{E}),
    \end{cases}
\end{equation}
and the acceleration of mass particle $\vec{a}$ will satisfy
\begin{equation}\label{lorentz}
    \vec{a}=-\vec{E}-4\vec{v}\times\vec{B}.
\end{equation}

Since \eqref{maxwell_0} are identical to Maxwell's equations, and \eqref{lorentz} is identical to the Lorentz force equation except for an overall minus sign and a factor of $4$ in the ``magnetic force'' term, we assume that the constants $\varepsilon_{\text{G}0}$ and $\mu_{\text{G}0}$ can simply vary as constants $\varepsilon_{0}$ and $\mu_{0}$ in electromagnetic theories, which means, \eqref{lorentz} is still valid and \eqref{maxwell_0} becomes
\begin{equation}\label{maxwell}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=\vec{j}+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
where $\vec{D}=\varepsilon_{\text{G}}\vec{E}$, $\vec{B}=\mu_{\text{G}}\vec{H}$. Also, we define $c=\frac{1}{\sqrt{\varepsilon_{\text{G}}\mu_{\text{G}}}}$, $G=\frac{1}{4\pi\varepsilon_{\text{G}}}$, together with $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\bar{h}_{0 \mu}=-4A_\mu/c=-4(\varphi/c,\vec{A})/c$.

We assume that the source of GW is located in a spherical coordinate origin and $c=c(r)$, $G=G(r)$, We suppose that $\bar{h}_{0 \mu}(\rho,\vec{j};c_0,G_0)$ is the solution of $\bar{h}_{0 \mu}$ when $\rho$ and $\vec{j}$ are assumed and $c\equiv c_0$, $G\equiv G_0$ everywhere, then if $c\equiv c_\text{s}$, $G\equiv G_\text{s}$, where $c_\text{s}$ and $G_\text{s}$ are two constants, within a region $r\leq R$, then $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be a solution of $\bar{h}_{0 \mu}$ within the region $r\leq R$, and now we can solve the equations
\begin{equation}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=0,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
with the boundary conditions $\vec{E}|_{r=R}=\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$, $\vec{B}|_{r=R}=\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$, where $\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$ and $\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be derived from $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$.

We can use the Li\'enard--Wiechert potentials formula to calculate the $\vec{E}$ and $\vec{B}$ field produced by a moving particle, whose position is $\vec{r}'$, in vacuum, and the results are
\begin{equation}
    \vec{E}=\frac{m}{4 \pi \varepsilon_{\text{G}0}}\left[\frac{(\vec{n}'-\vec{\beta})}{(1-\vec{\beta} \cdot \vec{n}')^{3} \gamma^{2} {r'}^{2}}+\frac{\vec{n}' \times\{(\vec{n}'-\vec{\beta}) \times \dot{\vec{\beta}}\}}{(1-\vec{\beta} \cdot \vec{n}')^{3} c_0 r'}\right]_{\text{ret}},\vec{B}=\left[\frac{\vec{n}'\times \vec{E}}{c_0}\right]_{\text{ret}},
\end{equation}
where $r'=\lvert\vec{r}'\rvert$, $\vec{n}'=\vec{r}'/r'$, $\vec{\beta}=\dot{\vec{r}}'/c_0$, $\gamma=1/\sqrt{1-\lvert\vec{\beta}\rvert^2}$, and ``ret'' denotes that the fields are ``retarded fields''.
\end{document}