### 2022-12-14 paper.tex
r part

\documentclass{article}
\usepackage{amsmath}
\def\d{\mathrm{d}}
\def\p{\partial}
\begin{document}
If we only take the lowest order effects of the motion of mass sources into account and neglect stresses, then we can define two vector fields, $\vec{E}$ and $\vec{B}$, which are similar to the electric and magnetic field. We can define $A_\mu=-\frac{1}{4}c_0\bar{h}_{0 \mu}=(\varphi/c_0,\vec{A})$, then $A_\mu$, which are similar to the electromagnetic 4-potential, will satisfy
\begin{equation}
    \p^\nu\p_\nu A_\mu=-\frac{4\pi G_0}{c_0^2}J_\mu,\quad\p^\mu A_\mu=0,
\end{equation}
where $J_\mu=-c_0T_{0 \mu}/c_0^2=(c_0\rho,\vec{j})=\rho(c_0,\vec{v})$ is 4-momentum density. If we define $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\varepsilon_{\text{G}0}=\frac{1}{4\pi G_0}$, $\mu_{\text{G}0}=\frac{4\pi G_0}{c_0^2}$, then there will be
\begin{equation}\label{maxwell_0}
    \begin{cases}
        \vec{\nabla}\cdot(\varepsilon_{\text{G}0}\vec{E})=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times(\mu_{\text{G}0}^{-1}\vec{B})=\vec{j}+\frac{\p}{\p t}(\varepsilon_{\text{G}0}\vec{E}),
    \end{cases}
\end{equation}
and the acceleration of mass particle $\vec{a}$ will satisfy
\begin{equation}\label{lorentz}
    \vec{a}=-\vec{E}-4\vec{v}\times\vec{B}.
\end{equation}

Since \eqref{maxwell_0} are identical to Maxwell's equations, and \eqref{lorentz} is identical to the Lorentz force equation except for an overall minus sign and a factor of $4$ in the ``magnetic force'' term, we assume that the constants $\varepsilon_{\text{G}0}$ and $\mu_{\text{G}0}$ can simply vary as constants $\varepsilon_{0}$ and $\mu_{0}$ in electromagnetic theories, which means, \eqref{lorentz} is still valid and \eqref{maxwell_0} becomes
\begin{equation}\label{maxwell}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=\vec{j}+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
where $\vec{D}=\varepsilon_{\text{G}}\vec{E}$, $\vec{B}=\mu_{\text{G}}\vec{H}$. Also, we define $c=\frac{1}{\sqrt{\varepsilon_{\text{G}}\mu_{\text{G}}}}$, $G=\frac{1}{4\pi\varepsilon_{\text{G}}}$, together with $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\bar{h}_{0 \mu}=-4A_\mu/c=-4(\varphi/c,\vec{A})/c$.

We assume that the source of GW is located in a spherical coordinate origin and $c=c(r)$, $G=G(r)$, We suppose that $\bar{h}_{0 \mu}(\rho,\vec{j};c_0,G_0)$ is the solution of $\bar{h}_{0 \mu}$ when $\rho$ and $\vec{j}$ are assumed and $c\equiv c_0$, $G\equiv G_0$ everywhere, then if $c\equiv c_\text{s}$, $G\equiv G_\text{s}$, where $c_\text{s}$ and $G_\text{s}$ are two constants, within a region $r\leq R$, then $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be a solution of $\bar{h}_{0 \mu}$ within the region $r\leq R$, and now we can solve the equations
\begin{equation}\label{maxwell_sf}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=0,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
with the boundary conditions $\vec{E}|_{r=R}=\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$, $\vec{B}|_{r=R}=\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$, where $\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$ and $\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be derived from $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$.

Again, since \eqref{maxwell_0} are identical to Maxwell's equations, we can use the Li\'enard--Wiechert potentials formula to calculate the $\vec{E}$ and $\vec{B}$ field produced by a moving particle, whose position is $\vec{r}'$, in vacuum, and the results are
\begin{equation}\label{lw}
    \vec{E}=\frac{m}{4 \pi \varepsilon_{\text{G}0}}\left[\frac{(\vec{n}'-\vec{\beta})}{(1-\vec{\beta} \cdot \vec{n}')^{3} \gamma^{2} {r'}^{2}}+\frac{\vec{n}' \times\{(\vec{n}'-\vec{\beta}) \times \dot{\vec{\beta}}\}}{(1-\vec{\beta} \cdot \vec{n}')^{3} c_0 r'}\right]_{\text{ret}},\vec{B}=\left[\frac{\vec{n}'\times \vec{E}}{c_0}\right]_{\text{ret}},
\end{equation}
where $r'=\lvert\vec{r}'\rvert$, $\vec{n}'=\vec{r}'/r'$, $\vec{\beta}=\dot{\vec{r}}'/c_0$, $\gamma=1/\sqrt{1-\lvert\vec{\beta}\rvert^2}$, and ``ret'' denotes that the fields are ``retarded fields''. Both $\vec{E}$ and $\vec{B}$ in \eqref{lw} can be divided into two parts: one, called ``inherent field part'', is proportional to $1/r'^2$ and the other one, called ``radiation part'', is proportional to $1/r'$. We only take the lowest order effects of the motion of mass sources into account, therefore we can assume that $\lvert\vec{\beta}\rvert\ll 1$. The ``inherent field part'' of $\vec{E}$ will be parallel to $\vec{r}'$ and the ``inherent field part'' of $\vec{B}$ will be nearly zero vector. The ``radiation part'' of $\vec{E}$ and $\vec{B}$ will be perpendicular to $\vec{r}'$. Since \eqref{maxwell_sf} are linear equations, we can discuss two parts of $\vec{E}$ and $\vec{B}$ separately for a binary GW source. If we assume that the distance between two components of binary is much less than $R$ mentioned above, then the ``inherent field part'' of $\vec{E}$ will have only $r$ component, and the ``radiation part'' of $\vec{E}$ and $\vec{B}$ will have no $r$ component.

For the ``inherent field part'', \eqref{maxwell_sf} will become $\frac{1}{r^2}\frac{\p}{\p r}(r^2 D_r)=0$, then $D_r\propto 1/r^2$, and $E_r\propto \varepsilon_{\text{G}}^{-1}(1/r^2)$. As mentioned below, for the ``radiation part'', $E\propto c^{-1}\varepsilon_{\text{G}}^{-1/2}(1/r)$, therefore the ``inherent field part'' will vanish in the region far away from source.

For the ``radiation part'', from the third and fourth equation of \eqref{maxwell_sf}, we can derive
\begin{equation}
    \begin{cases}
        [\vec{e}_r\cdot(\vec{\nabla}\times\vec{E})]\vec{e}_r-\frac{1}{r}\frac{\p}{\p r}(r{E}_\phi)\vec{e}_\theta+\frac{1}{r}\frac{\p}{\p r}(r{E}_\theta)\vec{e}_\phi=-\mu_{\text{G}}\frac{\p}{\p t}({H}_\theta\vec{e}_\theta+{H}_\phi\vec{e}_\phi),\\
        [\vec{e}_r\cdot(\vec{\nabla}\times\vec{H})]\vec{e}_r-\frac{1}{r}\frac{\p}{\p r}(r{H}_\phi)\vec{e}_\theta+\frac{1}{r}\frac{\p}{\p r}(r{H}_\theta)\vec{e}_\phi=+\varepsilon_{\text{G}}\frac{\p}{\p t}({E}_\theta\vec{e}_\theta+{E}_\phi\vec{e}_\phi),
    \end{cases}
\end{equation}
and then
\begin{equation}
    \begin{cases}
        \frac{1}{r}\frac{\p}{\p r}(r{E}_\theta)
        =-\mu_{\text{G}}\frac{\p}{\p t}{H}_\phi,\\
        \frac{1}{r}\frac{\p}{\p r}(r{E}_\phi)
        =+\mu_{\text{G}}\frac{\p}{\p t}{H}_\theta,\\
        \frac{1}{r}\frac{\p}{\p r}(r{H}_\theta)
        =+\varepsilon_{\text{G}}\frac{\p}{\p t}{E}_\phi,\\
        \frac{1}{r}\frac{\p}{\p r}(r{H}_\phi)
        =-\varepsilon_{\text{G}}\frac{\p}{\p t}{E}_\theta,\\
    \end{cases}
\end{equation}
\begin{equation}
    \begin{cases}
        \frac{\p}{\p r}(r{E}_\theta)
        =-\mu_{\text{G}}\frac{\p}{\p t}(r{H}_\phi),\\
        \frac{\p}{\p r}(r{E}_\phi)
        =+\mu_{\text{G}}\frac{\p}{\p t}(r{H}_\theta),\\
        \frac{\p}{\p r}(r{H}_\theta)
        =+\varepsilon_{\text{G}}\frac{\p}{\p t}(r{E}_\phi),\\
        \frac{\p}{\p r}(r{H}_\phi)
        =-\varepsilon_{\text{G}}\frac{\p}{\p t}(r{E}_\theta),\\
    \end{cases}
\end{equation}
\begin{equation}
    \begin{cases}
        \mu_{\text{G}}\frac{\p}{\p r}[\mu_{\text{G}}^{-1}\frac{\p}{\p r}(r{E}_\theta)]
        =-\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{H}_\phi)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{E}_\theta),\\
        \mu_{\text{G}}\frac{\p}{\p r}[\mu_{\text{G}}^{-1}\frac{\p}{\p r}(r{E}_\phi)]
        =+\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{H}_\theta)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{E}_\phi),\\
        \varepsilon_{\text{G}}\frac{\p}{\p r}[\varepsilon_{\text{G}}^{-1}\frac{\p}{\p r}(r{H}_\theta)]
        =+\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{E}_\phi)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{H}_\theta),\\
        \varepsilon_{\text{G}}\frac{\p}{\p r}[\varepsilon_{\text{G}}^{-1}\frac{\p}{\p r}(r{H}_\phi)]
        =-\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{E}_\theta)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{H}_\phi),\\
    \end{cases}
\end{equation}
therefore
\begin{equation}\label{radiation_part}
    \begin{cases}
        \frac{\p}{\p r}\frac{\p}{\p r}(r{E}_\theta)-\frac{\p}{\p r}(\ln\mu_{\text{G}})\frac{\p}{\p r}(r{E}_\theta)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{E}_\theta)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{E}_\phi)-\frac{\p}{\p r}(\ln\mu_{\text{G}})\frac{\p}{\p r}(r{E}_\phi)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{E}_\phi)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{H}_\theta)-\frac{\p}{\p r}(\ln\varepsilon_{\text{G}})\frac{\p}{\p r}(r{H}_\theta)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{H}_\theta)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{H}_\phi)-\frac{\p}{\p r}(\ln\varepsilon_{\text{G}})\frac{\p}{\p r}(r{H}_\phi)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{H}_\phi)=0.\\
    \end{cases}
\end{equation}
All equations in \eqref{radiation_part} have the form like
\begin{equation}\label{radiation}
    \frac{\p^2}{\p r^2}f(r,t)-p(r)\frac{\p}{\p r}f(r,t)-\frac{\p^2}{\p (ct)^2}f(r,t)=0.
\end{equation}
If we solve \eqref{radiation} by the method of separation of variables, we can suppose that $f(r,t)=f(r)e^{-ikct}$, and then
\begin{equation}\label{radiation_r}
    \frac{\d^2}{\d r^2}f(r)-p(r)\frac{\d}{\d r}f(r)+k^2f(r)=0.
\end{equation}
If $p$ is constant, then
\begin{equation}
    f(r)=e^{(p/2)r}[C_+e^{i\sqrt{k-(p/2)^2}r}+C_-e^{-i\sqrt{k-(p/2)^2}r}],
\end{equation}
where $C_+$ and $C_-$ are two constants, and
\begin{equation}
    f(r,t)=e^{(p/2)r}[C_+e^{i(+\sqrt{(\omega/c)-(p/2)^2}r-\omega t)}+C_-e^{i(-\sqrt{(\omega/c)-(p/2)^2}r-\omega t)}],
\end{equation}
where $\omega=kc$. If $\d p/\d r$ is small, we can divide the region $r> R$ into many spherical shells, and in each of them $p$ is nearly constant. To make $f(r,t)$ continuous anytime, we suppose that the ``angular frequency'' $\omega$ in each shell is same, and then there will be a approximate solution
\begin{equation}
    f(r,t)=e^{\int(p/2)\d r}[C_+e^{i(+\int\sqrt{(\omega/c)-(p/2)^2}\d r-\omega t)}+C_-e^{i(-\int\sqrt{(\omega/c)-(p/2)^2}\d r-\omega t)}].
\end{equation}
\end{document}