### 2022-12-15 paper.tex
methods ok

\documentclass{article}
\usepackage{amsmath}
\def\d{\mathrm{d}}
\def\p{\partial}
\begin{document}
If we only take the lowest order effects of the motion of mass sources into account and neglect stresses, then we can define two vector fields, $\vec{E}$ and $\vec{B}$, which are similar to the electric and magnetic field. We can define $A_\mu=-\frac{1}{4}c_0\bar{h}_{0 \mu}=(\varphi/c_0,\vec{A})$, then $A_\mu$, which are similar to the electromagnetic 4-potential, will satisfy
\begin{equation}
    \p^\nu\p_\nu A_\mu=-\frac{4\pi G_0}{c_0^2}J_\mu,\quad\p^\mu A_\mu=0,
\end{equation}
where $J_\mu=-c_0T_{0 \mu}/c_0^2=(c_0\rho,\vec{j})=\rho(c_0,\vec{v})$ is 4-momentum density. If we define $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\varepsilon_{\text{G}0}=\frac{1}{4\pi G_0}$, $\mu_{\text{G}0}=\frac{4\pi G_0}{c_0^2}$, then there will be
\begin{equation}\label{maxwell_0}
    \begin{cases}
        \vec{\nabla}\cdot(\varepsilon_{\text{G}0}\vec{E})=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times(\mu_{\text{G}0}^{-1}\vec{B})=\vec{j}+\frac{\p}{\p t}(\varepsilon_{\text{G}0}\vec{E}),
    \end{cases}
\end{equation}
and the acceleration of mass particle $\vec{a}$ will satisfy
\begin{equation}\label{lorentz}
    \vec{a}=-\vec{E}-4\vec{v}\times\vec{B}.
\end{equation}

Since \eqref{maxwell_0} are identical to Maxwell's equations, and \eqref{lorentz} is identical to the Lorentz force equation except for an overall minus sign and a factor of $4$ in the ``magnetic force'' term, we assume that the constants $\varepsilon_{\text{G}0}$ and $\mu_{\text{G}0}$ can simply vary as constants $\varepsilon_{0}$ and $\mu_{0}$ in electromagnetic theories, which means, \eqref{lorentz} is still valid and \eqref{maxwell_0} becomes
\begin{equation}\label{maxwell}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=\rho,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=\vec{j}+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
where $\vec{D}=\varepsilon_{\text{G}}\vec{E}$, $\vec{B}=\mu_{\text{G}}\vec{H}$. Also, we define $c=\frac{1}{\sqrt{\varepsilon_{\text{G}}\mu_{\text{G}}}}$, $G=\frac{1}{4\pi\varepsilon_{\text{G}}}$, together with $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $\vec{B}=\vec{\nabla}\times\vec{A}$, and $\bar{h}_{0 \mu}=-4A_\mu/c=-4(\varphi/c,\vec{A})/c$.

We assume that the source of GW is located in a spherical coordinate origin and $c=c(r)$, $G=G(r)$, We suppose that $\bar{h}_{0 \mu}(\rho,\vec{j};c_0,G_0)$ is the solution of $\bar{h}_{0 \mu}$ when $\rho$ and $\vec{j}$ are assumed and $c\equiv c_0$, $G\equiv G_0$ everywhere, then if $c\equiv c_\text{s}$, $G\equiv G_\text{s}$, where $c_\text{s}$ and $G_\text{s}$ are two constants, within a region $r\leq R$, then $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be a solution of $\bar{h}_{0 \mu}$ within the region $r\leq R$, and now we can solve the equations
\begin{equation}\label{maxwell_sf}
    \begin{cases}
        \vec{\nabla}\cdot\vec{D}=0,\\
        \vec{\nabla}\cdot\vec{B}=0,\\
        \vec{\nabla}\times\vec{E}=-\frac{\p}{\p t}\vec{B},\\
        \vec{\nabla}\times\vec{H}=+\frac{\p}{\p t}\vec{D},
    \end{cases}
\end{equation}
with the boundary conditions $\vec{E}|_{r=R}=\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$, $\vec{B}|_{r=R}=\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$, where $\vec{E}(\rho,\vec{j};c_\text{s},G_\text{s})$ and $\vec{B}(\rho,\vec{j};c_\text{s},G_\text{s})$ can be derived from $\bar{h}_{0 \mu}(\rho,\vec{j};c_\text{s},G_\text{s})$.

Again, since \eqref{maxwell_0} are identical to Maxwell's equations, we can use the Li\'enard--Wiechert potentials formula to calculate the $\vec{E}$ and $\vec{B}$ field produced by a moving particle, whose position is $\vec{r}'$, in vacuum, and the results are
\begin{equation}\label{lw}
    \vec{E}=\frac{m}{4 \pi \varepsilon_{\text{G}0}}\left[\frac{(\vec{n}'-\vec{\beta})}{(1-\vec{\beta} \cdot \vec{n}')^{3} \gamma^{2} {r'}^{2}}+\frac{\vec{n}' \times\{(\vec{n}'-\vec{\beta}) \times \dot{\vec{\beta}}\}}{(1-\vec{\beta} \cdot \vec{n}')^{3} c_0 r'}\right]_{\text{ret}},\vec{B}=\left[\frac{\vec{n}'\times \vec{E}}{c_0}\right]_{\text{ret}},
\end{equation}
where $r'=\lvert\vec{r}'\rvert$, $\vec{n}'=\vec{r}'/r'$, $\vec{\beta}=\dot{\vec{r}}'/c_0$, $\gamma=1/\sqrt{1-\lvert\vec{\beta}\rvert^2}$, and ``ret'' denotes that the fields are ``retarded fields''. Both $\vec{E}$ and $\vec{B}$ in \eqref{lw} can be divided into two parts: one, called ``inherent field part'', is proportional to $1/r'^2$ and the other one, called ``radiation part'', is proportional to $1/r'$. We only take the lowest order effects of the motion of mass sources into account, therefore we can assume that $\lvert\vec{\beta}\rvert\ll 1$. The ``inherent field part'' of $\vec{E}$ will be parallel to $\vec{r}'$ and the ``inherent field part'' of $\vec{B}$ will be nearly zero vector. The ``radiation part'' of $\vec{E}$ and $\vec{B}$ will be perpendicular to $\vec{r}'$. Since \eqref{maxwell_sf} are linear equations, we can discuss two parts of $\vec{E}$ and $\vec{B}$ separately for a binary GW source. If we assume that the distance between two components of binary is much less than $R$ mentioned above, then the ``inherent field part'' of $\vec{E}$ will have only $r$ component, and the ``radiation part'' of $\vec{E}$ and $\vec{B}$ will have no $r$ component.

For the ``inherent field part'', \eqref{maxwell_sf} will become $\frac{1}{r^2}\frac{\p}{\p r}(r^2 D_r)=0$, then $D_r\propto 1/r^2$, and $E_r\propto \varepsilon_{\text{G}}^{-1}(1/r^2)$. As mentioned below, for the ``radiation part'', $E\propto c^{-1}\varepsilon_{\text{G}}^{-1/2}(1/r)$, therefore the ``inherent field part'' will vanish in the region far away from source.

For the ``radiation part'', from the third and fourth equation of \eqref{maxwell_sf}, we can derive
\begin{equation}
    \begin{cases}
        [\vec{e}_r\cdot(\vec{\nabla}\times\vec{E})]\vec{e}_r-\frac{1}{r}\frac{\p}{\p r}(r{E}_\phi)\vec{e}_\theta+\frac{1}{r}\frac{\p}{\p r}(r{E}_\theta)\vec{e}_\phi=-\mu_{\text{G}}\frac{\p}{\p t}({H}_\theta\vec{e}_\theta+{H}_\phi\vec{e}_\phi),\\
        [\vec{e}_r\cdot(\vec{\nabla}\times\vec{H})]\vec{e}_r-\frac{1}{r}\frac{\p}{\p r}(r{H}_\phi)\vec{e}_\theta+\frac{1}{r}\frac{\p}{\p r}(r{H}_\theta)\vec{e}_\phi=+\varepsilon_{\text{G}}\frac{\p}{\p t}({E}_\theta\vec{e}_\theta+{E}_\phi\vec{e}_\phi),
    \end{cases}
\end{equation}
and then
\begin{equation}
    \begin{cases}
        \frac{1}{r}\frac{\p}{\p r}(r{E}_\theta)
        =-\mu_{\text{G}}\frac{\p}{\p t}{H}_\phi,\\
        \frac{1}{r}\frac{\p}{\p r}(r{E}_\phi)
        =+\mu_{\text{G}}\frac{\p}{\p t}{H}_\theta,\\
        \frac{1}{r}\frac{\p}{\p r}(r{H}_\theta)
        =+\varepsilon_{\text{G}}\frac{\p}{\p t}{E}_\phi,\\
        \frac{1}{r}\frac{\p}{\p r}(r{H}_\phi)
        =-\varepsilon_{\text{G}}\frac{\p}{\p t}{E}_\theta,\\
    \end{cases}
\end{equation}
\begin{equation}
    \begin{cases}
        \frac{\p}{\p r}(r{E}_\theta)
        =-\mu_{\text{G}}\frac{\p}{\p t}(r{H}_\phi),\\
        \frac{\p}{\p r}(r{E}_\phi)
        =+\mu_{\text{G}}\frac{\p}{\p t}(r{H}_\theta),\\
        \frac{\p}{\p r}(r{H}_\theta)
        =+\varepsilon_{\text{G}}\frac{\p}{\p t}(r{E}_\phi),\\
        \frac{\p}{\p r}(r{H}_\phi)
        =-\varepsilon_{\text{G}}\frac{\p}{\p t}(r{E}_\theta),\\
    \end{cases}
\end{equation}
\begin{equation}
    \begin{cases}
        \mu_{\text{G}}\frac{\p}{\p r}[\mu_{\text{G}}^{-1}\frac{\p}{\p r}(r{E}_\theta)]
        =-\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{H}_\phi)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{E}_\theta),\\
        \mu_{\text{G}}\frac{\p}{\p r}[\mu_{\text{G}}^{-1}\frac{\p}{\p r}(r{E}_\phi)]
        =+\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{H}_\theta)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{E}_\phi),\\
        \varepsilon_{\text{G}}\frac{\p}{\p r}[\varepsilon_{\text{G}}^{-1}\frac{\p}{\p r}(r{H}_\theta)]
        =+\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{E}_\phi)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{H}_\theta),\\
        \varepsilon_{\text{G}}\frac{\p}{\p r}[\varepsilon_{\text{G}}^{-1}\frac{\p}{\p r}(r{H}_\phi)]
        =-\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p r}(r{E}_\theta)
        =\varepsilon_{\text{G}}\mu_{\text{G}}\frac{\p}{\p t}\frac{\p}{\p t}(r{H}_\phi),\\
    \end{cases}
\end{equation}
therefore
\begin{equation}\label{radiation_part}
    \begin{cases}
        \frac{\p}{\p r}\frac{\p}{\p r}(r{E}_\theta)-\frac{\p}{\p r}(\ln\mu_{\text{G}})\frac{\p}{\p r}(r{E}_\theta)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{E}_\theta)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{E}_\phi)-\frac{\p}{\p r}(\ln\mu_{\text{G}})\frac{\p}{\p r}(r{E}_\phi)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{E}_\phi)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{H}_\theta)-\frac{\p}{\p r}(\ln\varepsilon_{\text{G}})\frac{\p}{\p r}(r{H}_\theta)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{H}_\theta)=0,\\
        \frac{\p}{\p r}\frac{\p}{\p r}(r{H}_\phi)-\frac{\p}{\p r}(\ln\varepsilon_{\text{G}})\frac{\p}{\p r}(r{H}_\phi)-\frac{\p}{\p (ct)}\frac{\p}{\p (ct)}(r{H}_\phi)=0.\\
    \end{cases}
\end{equation}
All equations in \eqref{radiation_part} have the form like
\begin{equation}\label{radiation}
    \frac{\p^2}{\p r^2}f(r,t)-p(r)\frac{\p}{\p r}f(r,t)-\frac{\p^2}{\p (ct)^2}f(r,t)=0.
\end{equation}
If we solve \eqref{radiation} by the method of separation of variables, we can suppose that $f(r,t)=f(r)e^{-ikct}$, and then
\begin{equation}\label{radiation_r}
    \frac{\d^2}{\d r^2}f(r)-p(r)\frac{\d}{\d r}f(r)+k^2f(r)=0.
\end{equation}
If $p$ is constant, then
\begin{equation}
    f(r)=e^{(p/2)r}[C_+e^{i\sqrt{k^2-(p/2)^2}r}+C_-e^{-i\sqrt{k^2-(p/2)^2}r}],
\end{equation}
where $C_+$ and $C_-$ are two constants, and
\begin{equation}
    f(r,t)=e^{(p/2)r}[C_+e^{i(+\sqrt{(\omega/c)^2-(p/2)^2}r-\omega t)}+C_-e^{i(-\sqrt{(\omega/c)^2-(p/2)^2}r-\omega t)}],
\end{equation}
where $\omega=kc$. If $\d p/\d r$ is small, we can divide the region $r> R$ into many spherical shells, and in each of them $p$ is nearly constant. To make $f(r,t)$ continuous anytime, we suppose that the ``angular frequency'' $\omega$ in each shell is the same, and then there will be a approximate solution
\begin{equation}\label{wave_solution}
    f(r,t)=e^{\int(p/2)\d r}[C_+e^{i(+\int\sqrt{(\omega/c)^2-(p/2)^2}\d r-\omega t)}+C_-e^{i(-\int\sqrt{(\omega/c)^2-(p/2)^2}\d r-\omega t)}],
\end{equation}
therefore
\begin{equation}
    f(r,t)=e^{\int O(p)\d r}[C_+e^{i(+\int[(\omega/c)^2+O(p^2)]\d r-\omega t)}+C_-e^{i(+\int[(\omega/c)^2+O(p^2)]\d r-\omega t)}].
\end{equation}
If $p$ is small, \eqref{wave_solution} can be approximated to
\begin{equation}
    f(r,t)=e^{\int(p/2)\d r}[C_+e^{i(+\int(\omega/c)\d r-\omega t)}+C_-e^{i(-\int(\omega/c)\d r-\omega t)}],
\end{equation}
then wave of $\vec{E}$ and $\vec{H}$ will approximately arrive Earth at the same time, and also $rE \propto e^{\int(\frac{\d}{\d r}(\ln\mu_{\text{G}})/2)\d r} \propto \mu_{\text{G}}^{1/2}$, $rH\propto e^{\int(\frac{\d}{\d r}(\ln\varepsilon_{\text{G}})/2)\d r} \propto \varepsilon_{\text{G}}^{1/2}$. Again, if $p$ is small, $\left\lvert i(\omega/c)\int f\d r\right\rvert \approx\left\lvert \frac{\p}{\p r}\int f\d r\right\rvert $ and $-i\omega\int f\d t=\frac{\p}{\p t}\int f\d t$. Since $\vec{H}=\mu_\text{G}^{-1}\vec{\nabla}\times\vec{A}$ and $H\propto \varepsilon_\text{G}^{1/2}/r$, $\mu_\text{G}^{-1}(\omega/c)A\propto\varepsilon_{\text{G}}^{1/2}/r\Rightarrow A\propto \mu_\text{G}^{1/2}/r$. Since $\vec{E}=-\vec{\nabla}\varphi-\frac{\p}{\p t}\vec{A}$, $E\propto \mu_\text{G}^{1/2}/r$ and $\frac{\p}{\p t}\vec{A}\propto \omega\mu_\text{G}^{1/2}/r\propto \mu_\text{G}^{1/2}/r$, $(\omega/c)\varphi\propto\mu_\text{G}^{1/2}/r\Rightarrow \varphi/c\propto\mu_\text{G}^{1/2}/r$. Therefore, $\bar{h}_{0 \mu}=-4(\varphi/c,\vec{A})/c\propto\mu_\text{G}^{1/2}/cr\propto G^{1/2}/c^2r$.

It is known that if $c\equiv c_0$, $G\equiv G_0$ everywhere, then the amplitude $h$ of GW produced by an inspiring binary satisfies
\begin{equation}
    h\propto\frac{\frac{G_0\mathcal{M}}{c_0^3}[\pi \frac{G_0\mathcal{M}}{c_0^3}F(t)]^{2/3}}{r/c_0}\cos[\int 2\pi F(t)\,\d t],
\end{equation}
where $\mathcal{M}$ is the chirp mass of binary and $F$ is the frequency of GW. Therefore in the region $r\leq R$ mentioned above,
\begin{equation}
    h\propto\frac{\frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}[\pi \frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}F(t)]^{2/3}}{r/c_\text{s}}\cos[\int 2\pi F(t)\,\d t],
\end{equation}
and then on Earth,
\begin{equation}
    h\propto\sqrt{\frac{c_\text{s}^4/G_\text{s}}{c_0^4/G_0}}\frac{\frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}[\pi \frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}F(t)]^{2/3}}{d/c_\text{s}}\cos[\int 2\pi F(t)\,\d t],
\end{equation}
where $d$ is the physical distance of binary. Considering the red shift, the amplitude of GW produced by an inspiring binary on Earth satisfies
\begin{equation}
    h\propto\sqrt{\frac{c_\text{s}^4/G_\text{s}}{c_0^4/G_0}}\frac{\frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}[\pi \frac{G_\text{s}\mathcal{M}}{c_\text{s}^3}F(t)]^{2/3}}{d/c_\text{s}}\cos[\int 2\pi \frac{F(t)}{1+z}\,\d t],
\end{equation}
where $z$ is the red shift of binary. If we define $F_\text{obs}(t)=F(t)/(1+z)$ and $\mathcal{M}_\text{obs}=\mathcal{M}(1+z)$, then
\begin{equation}
    h\propto\sqrt{\frac{c_\text{s}^6/G_\text{s}}{c_0^6/G_0}}\frac{\frac{G_\text{s}\mathcal{M}_\text{obs}}{c_\text{s}^3}[\pi \frac{G_\text{s}\mathcal{M}_\text{obs}}{c_\text{s}^3}F_\text{obs}(t)]^{2/3}}{d_\text{L}/c_0}\cos[\int 2\pi F_\text{obs}(t)\,\d t],
\end{equation}
where $d_\text{L}=d(1+z)$ is the luminosity distance of binary. However, in former researches, $h$ was assumed to satisfy
\begin{equation}
    h\propto\frac{\frac{G_0\mathcal{M}_\text{obs}}{c_0^3}[\pi \frac{G_0\mathcal{M}_\text{obs}}{c_0^3}F_\text{obs}(t)]^{2/3}}{d_\text{L}/c_0}\cos[\int 2\pi F_\text{obs}(t)\,\d t],
\end{equation}
therefore, $d_\text{L,G}$, the result of measuring $d_\text{L}$ by using GW observation data and methods in former researches, will satisfy
\begin{equation}
    \sqrt{\frac{G_\text{s}/c_\text{s}^6}{G_0/c_0^6}}=\frac{d_\text{L,G}}{d_\text{L}},
\end{equation}
where $d_\text{L}$ can be also measured precisely by using EMW observation data.

Therefore, we can constrain $c_\text{s}$ and $G_\text{s}$ by comparing $d_\text{L}$ and $d_\text{L,G}$, and $\sqrt{[\frac{1}{2}\Delta\!\ln(G_\text{s}/G_0)]^2+[3\Delta\!\ln(c_\text{s}/c_0)]^2}=\Delta\!\ln(d_\text{L,G}/d_\text{L})$, where $\Delta$ denotes to the uncertainty. The result of measuring $d_\text{L}$ and $d_\text{L,G}$ for GW170817/GRB 170817A is $d_\text{L}=(42.9 \pm 3.2)\text{Mpc}$ and $d_\text{L,G}=40_{-14}^{+8}\text{Mpc}$, and we estimate that this result can constrain $c_\text{s}$ and $G_\text{s}$ as $(G_\text{s}/G_0)^{1/2}(c_\text{s}/c_0)^3=0.9 \pm 0.3$.
\end{document}